### 2015-04-26 evaluation.tex
various updates

We can see that \whoop reports many more races than \corral. This is expected, though, as \whoop over-approximates the original driver for soundness and scalability. This means that if a race exists, \whoop will report it (assuming tool and model correctness). On the other hand, \corral performs bounded checking, which means that it will report a race only under pre-defined bounds. A common problem for \corral when analyzing drivers is loop coverage. To deal with this we enable the over-approximation technique described in~\cite{lal2014powering}. This can potentially lead \corral to report false positives, but we have not seen this in practice.

We found the following data race in the generic\_nvram driver: during the \texttt{llseek} file operation, the driver is accessing the file offset (\texttt{file->f\_pos}) without using a lock. When we investigated another similar driver that uses the same API, we saw that it was protecting the same offset access with a mutex. This made us suspicious that the generic\_nvram driver might have a real race. We filed a bug report, but did not manage to get a response yet.

A common case of races that we discovered had to do with incrementing a counter without using a lock. This might happen for performance, but even benign races might lead to serious issues, especially if a compiler tries to perform optimizations assuming no races exist.

@@

In Table~\ref{tab:results}, we can notice that \whoop is much faster than \corral. For example, \whoop analyses the dtlk char driver in 6.51 seconds, whereas \corral requires 1287.05 seconds.

>>>>

We can see that, in general, \whoop reports (many) more races than \corral. This is expected, though, as \whoop over-approximates the original driver for soundness. This means that if a race exists, \whoop will report it (assuming tool and model correctness). On the other hand, \corral performs bounded checking, which means that it will report a race only under pre-defined bounds. It is worthwhile to note that a common problem for \corral when analyzing drivers is loop coverage. To deal with this we enable the over-approximation described in~\cite{lal2014powering}. This can potentially lead \corral to report false positives, but we have not seen this in practice.

Most of the data races that \whoop and \corral found can be classified in two cases. The first case had to do with incrementing a global counter without using a lock. The reason behind this might be performance, but even benign races can lead to undefined behavior according to the C standard. The second case had to do with an entry point modifying a field of an object (either global or passed as parameter) without using a lock. This can easily lead to a data race if another concurrently running entry point accesses the same field of the same object. As an example of the second case, we discovered the following data race in the generic\_nvram char driver: during the \texttt{llseek} entry point, the driver is accessing the file offset \texttt{file->f\_pos} (\texttt{file} is passed as a parameter to the entry point) without first acquiring a lock. This can lead to a race if the driver invokes \texttt{llseek} from another thread passing the same \texttt{file} object as a parameter. When we investigated another char driver that uses the same API, we saw that it was protecting the offset access in the same entry point with a mutex. This made us suspicious that the race we found in generic\_nvram is real. We filed a bug report, but did not manage to get a response yet.

@@

We can notice that \whoop is significantly faster than \corral. This is expected as \whoop achieves scalability using over-approximation and summarization techniques. This allows \whoop to perform especially well in complex drivers such as the r8169 ethernet driver: although \corral ... and timeouts with a csb of 9, \whoop manages to analyze the driver in 228.1 seconds. We believe that the reason behind this is that r8169 has many loops and uses deeply-nested recursion in some entry points, which \corral might not be able to handle efficiently.
